Physcraper: Body

### 2020-07-01 2-body.Rmd
cleaning comments

Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field. In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
Automated approaches often to suffer from errors in homology inference that are rapidly caught by human oversight. ()
Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships.  \comment(Citations... I think there are some Scott Edwards papers...https://www.pnas.org/content/109/37/14942, maybe Roajs et al...)
It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].

@@

- Technically, any DNA molecular database can be used to search for new sequences.
  By default we rely on the GenBank database.
  The new sequence search can be performed on the remote database or in a local database.
- The next step is to identify the search taxon within the reference taxonomy. The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group. Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy. The search taxon can be determined by the user. If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. \comment(hmm, a figure might be good for this. I can work on one!) This is known as the Most Recent Common
  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) \comment(prob need a cite here).
  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>). I FORGOT TO ASK YOU ABOUT THIS CITATION! \comment( it's here! https://bdj.pensoft.net/article/12581/instance/3545675/)

@@

- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
Longer sequences belonging to the
  same taxon as the original sequence will be considered further for analysis.
- Among the filtered sequences, there are often several representatives per taxon.
  Although it can be useful to keep some of them, for example, to investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
- Reverse complement sequences are identified and translated.
- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences. This can be done iteratively, but by default only sequences in the checked alignment are blasted. I ALSO FORGOT TO ASK: ***Is there an argument to control the number of cycles of blast searches with new sequences?*** \comment(there's only two options - one cycle or, -r where it runs until it doesn't find anything new)
- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.

@@

All these tools are key efforts for advancing towards reproducibility in phylogenetics,
a field that has relied on processes which are somewhat artisanal.
In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates. By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
community to build up our phylogenetic knowledge, we can use not only data accumulated in
molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
While not generating full phylogenies *de novo*, physcraper is still capable of generating new phylogenetic knowledge.
In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??). \comment(hmmmm mayeb cut bc I don't know!)
It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.

@@

The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
\comment(need to add teh OPenTree ABI grant..)
We acknowledge contributions from ...

>>>>

Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field. In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.].
It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].

@@

- Technically, any DNA molecular database can be used to search for new sequences.
  By default we rely on the GenBank database.
  The new sequence search can be performed on the remote database or in a local database.
- The next step is to identify the search taxon within the reference taxonomy. The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group. Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy. The search taxon can be determined by the user. If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool [Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].

@@

- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
Longer sequences belonging to the
  same taxon as the original sequence will be considered further for analysis.
- Among the filtered sequences, there are often several representatives per taxon.
  Although it can be useful to keep some of them, for example, to investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
- Reverse complement sequences are identified and translated.
- Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found. By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
- Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.

@@

All these tools are key efforts for advancing towards reproducibility in phylogenetics,
a field that has relied on processes which are somewhat artisanal.
In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates. By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
community to build up our phylogenetic knowledge, we can use not only data accumulated in
molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
While not generating full phylogenies *de novo*, physcraper is still capable of generating new phylogenetic knowledge.
It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.

@@

Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838.
Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No. ACI-1429783.

We thank member of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
